---
title: 'APIs'
author: "Claire Haldeman"
date: "2/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Application Programming Interfaces (APIs)

Why use an API?

1. Some things are only available through an API (e.g. Twitter)
2. It's more reproducible, everyone can access the data
3. It can scale much more easily than point-and-click downloads

Why build an API?

1. It allows you to scale your data
2. It promotes a standardized approach to data management
3. It democratizes access to data and methods--provides a way for people with lots of processing power to process for otehrs

Technical Context

* Server: the host
* Client : us
* HTTP: ettiquette for how clients and servers communicate
* Request: specific information about what you want, usually as a URL
* Response: what the server sends you back 
* URL: "Uniform resource locator" is a string that specifies a web location, a method for retrieving information, and other things

(Note: Census has great API and documentation)

How to access API

1. Build up a URL from API documentation 
  * begin query string with a ?
  * use "GET=" to specify which variables you would like it to return
  * use "&" to separate different parameters
  * use "*" to indicate "all" or "placeholder"
2. Check for errors
3. Import data from JSON
  * JSON is JavaScript Object N?
  * A hierarchical data structure
  * Usually we only care about a few of the levels of the hierarchy

How to use R to do this for you

```{r}
library(tidyverse)
library(httr)
library(jsonlite)
```


```{r, eval = FALSE}
#make url, specify outside of the call

api_url <- "http:theurl"

#use GET(api_url, user_agent = ) to make request, declare yourself (make sure you're adhering to terms of service)

#use http_status to check for errors

#download data with content( , as = text)

#read from JSON into matrix

#convert matrix into tbl
```

Example:

```{r}
#generic: "https://educationdata.urban.org/api/v1/{topic}/{source}/{endpoint}/{year}/[additional_specifiers _or_disaggregators]/[optional filters]"

api_url <- "https://educationdata.urban.org/api/v1/college-university/ipeds/enrollment-full-time-equivalent/2016/Graduate/?fips=11"

ed_json <- GET(url = api_url,
               user_agent("Brookings Researcher (chaldeman@brookings.edu)"))

http_status(ed_json)

ed_json <- content(ed_json, type = "text")

ed_json <- fromJSON(ed_json)

ed_data <- ed_json[["results"]]

```

Be aware: 
* sometimes you need to authenticate yourself
* sometimes you get data in multiple pages (see notes for help!)

But!! Most big APIs have an R package that helps you access it
